---
title: "Untitled"
author: "Christoph Kurz"
date: "May 30, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r cars}
library(rstan)
library(ggplot2)
library(flexmix)
library(countreg)
```

### DP-NB model definition in `stan`

```{r}
model_string <-"
data {
int<lower=1> K; // number of mixture components
int<lower=1> N; // number of data points
int<lower=1> M; //the number of columns in the model matrix
matrix[N,M] X; //the model matrix
int y[N]; // observations
real<lower=0> alpha0 ; // dirichlet prior
}
transformed data {
vector<lower=0>[K] alpha0_vec = rep_vector(alpha0, K); //symmetric dirichlet prior
}
parameters {
  positive_ordered[K] gamma; // primitives of mixing proportions
  vector[M] betas[K];
  real<lower=0> sigma[K]; // scales of mixture components
}
transformed parameters {
  vector[K] theta = gamma / sum(gamma);
}
model {
gamma ~ gamma(alpha0_vec, 1); // implies: theta ~ dirichlet(alpha_vec)
sigma ~ lognormal(0, 2);
for (k in 1:K)
betas[k,] ~ normal(0,10);
for (n in 1:N) {
real lps[K];
for (k in 1:K)
lps[k] = log(theta[k]) + neg_binomial_2_lpmf(y[n] | exp(X[n] * betas[k]), sigma[k]);
target += log_sum_exp(lps);
}
}"
```

### `flexmix` driver for Negative Binomial regression, adapted from the `countreg` package with improvements to the optimization part

```{r, echo=T}
FLXMRnegbin2 <- function (formula = . ~ ., theta = NULL, offset = NULL, control = list(reltol = .Machine$double.eps^(1/1.5), 
    maxit = 500)) 
{
    .theta <- theta
    nbrefit <- function(x, y, w) {
        fit <- c(glm.fit(x, y, weights = w, offset = offset, 
            family = MASS::negative.binomial(theta)), list(call = sys.call(), 
            offset = offset, control = eval(formals(glm.fit)$control), 
            method = "weighted.glm.fit"))
        fit$df.null <- sum(w) + fit$df.null - fit$df.residual - 
            fit$rank - is.null(.theta)
        fit$df.residual <- sum(w) - fit$rank - is.null(.theta)
        fit$x <- x
        fit
    }
    z <- methods::new("FLXMRglm", weighted = TRUE, formula = formula, 
        name = "FLXMRglm: negative.binomial", offset = offset, 
        family = "negative.binomial", refit = nbrefit)
    z@preproc.y <- function(x) {
        if (ncol(x) > 1L) 
            stop(paste("for the", family, "family y must be univariate"))
        x
    }
    z@defineComponent <- if (is.null(.theta)) {
        expression({
            predict <- function(x, ...) {
                dotarg <- list(...)
                if ("offset" %in% names(dotarg)) offset <- dotarg$offset
                p <- x %*% coef
                if (!is.null(offset)) p <- p + offset
                exp(p)
            }
            logLik <- function(x, y, ...) suppressWarnings(dnbinom(y, 
                mu = predict(x, ...), size = theta, log = TRUE))
            methods::new("FLXcomponent", parameters = list(coef = coef, 
                theta = theta), logLik = logLik, predict = predict, 
                df = df)
        })
    }
    else {
        as.expression(substitute({
            predict <- function(x, ...) {
                dotarg <- list(...)
                if ("offset" %in% names(dotarg)) offset <- dotarg$offset
                p <- x %*% coef
                if (!is.null(offset)) p <- p + offset
                exp(p)
            }
            logLik <- function(x, y, ...) suppressWarnings(dnbinom(y, 
                mu = predict(x, ...), size = theta, log = TRUE))
            methods::new("FLXcomponent", parameters = list(coef = coef), 
                logLik = logLik, predict = predict, df = df)
        }, as.environment(list(theta = .theta))))
    }
    z@fit <- function(x, y, w, component) {
        if (is.null(component$theta)) {
            df <- ncol(x)
            theta <- if (is.null(.theta)) 
                1
            else .theta
            cf <- glm.fit(x, y, weights = w, family = MASS::negative.binomial(theta), 
                offset = offset, start = component$coef)$coefficients
        }
        else {
            df <- ncol(x) + 1
            if (is.null(offset)) 
                offset <- 0
            nll <- function(par) {
                beta <- par[-df]
                theta <- exp(par[df])
                mu <- exp(drop(x %*% beta + offset))
                suppressWarnings(-sum(w * dnbinom(y, mu = mu, 
                  size = theta, log = TRUE)))
            }
            gr <- function(par) {
                beta <- par[-df]
                theta <- exp(par[df])
                mu <- exp(drop(x %*% beta + offset))
                gr <- drop(y - mu * (y + theta)/(mu + theta))
                colSums(-w * cbind(gr * x, theta * (digamma(y + 
                  theta) - digamma(theta) + log(theta) + 1 - 
                  log(mu + theta) - (y + theta)/(mu + theta))))
            }
            start <- c(component$coef, component$theta)
            if (length(start) < df) 
                start <- c(glm.fit(x, y, weights = w, family = MASS::negative.binomial(1), 
                  offset = offset)$coefficients, 0)
            opt <- optim(par = start, fn = nll, gr = gr, method = "L-BFGS-B", lower=0, upper=50,
                control = control)
            cf <- opt$par[-df]
            theta <- exp(opt$par[df])
        }
        with(list(coef = cf, theta = theta, df = ncol(x) + is.null(.theta)), 
            eval(z@defineComponent))
    }
    z
}

```

# Simulation Study

## 2 components
### heavy overlap

```{r}
set.seed(41)

N1 <- 3000
N2 <- 5000

X <- data.frame(V=rep(1,N1+N2), V1=runif(N1+N2, 0.8, 1), V2=runif(N1+N2, 0.8, 1))

y <- rnbinom(N1, size = 1.2, mu = exp(as.matrix(X) %*% c(1, -0.5, 0.4)))
y2 <- rnbinom(N2, size = 10, mu = exp(as.matrix(X) %*% c(1, 1.0, -0.5)))
```

density histogram:

```{r}
df <- data.frame(val = c(y,y2), comp = c(rep(1, length(y)), rep(2, length(y2))))
ggplot(df, aes(val, color=as.factor(comp))) + geom_density(adjust=1.5)
```

```{r, echo=F}
write.csv(df, "/home/ubu/ownCloud/docs/papers/paper1/data/2comp_heavy.csv", row.names = F)
```

compute Negative Binomial regression mixtures for 1-6 components

```{r, cache=T, warning=F}
comp2h1 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2), V1=X$V1, V2=X$V2), k = 1, model = FLXMRnegbin2())
comp2h2 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2), V1=X$V1, V2=X$V2), k = 2, model = FLXMRnegbin2())
comp2h3 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2), V1=X$V1, V2=X$V2), k = 3, model = FLXMRnegbin2())
comp2h4 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2), V1=X$V1, V2=X$V2), k = 4, model = FLXMRnegbin2())
comp2h5 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2), V1=X$V1, V2=X$V2), k = 5, model = FLXMRnegbin2())
comp2h6 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2), V1=X$V1, V2=X$V2), k = 6, model = FLXMRnegbin2())
```

compute AIC, BIC, and ICL of all models

```{r}
aics <- c(AIC(comp2h1), AIC(comp2h2), AIC(comp2h3), AIC(comp2h4), AIC(comp2h5), AIC(comp2h6))
bics <- c(BIC(comp2h1), BIC(comp2h2), BIC(comp2h3), BIC(comp2h4), BIC(comp2h5), BIC(comp2h6))
icls <- c(ICL(comp2h1), ICL(comp2h2), ICL(comp2h3), ICL(comp2h4), ICL(comp2h5), ICL(comp2h6))
```

find the best model

```{r}
which.min(aics)
which.min(bics)
which.min(icls)
```

compute DP-NB

```{r stan2h, warning=F, cache=T}
m2h <- stan(model_code = model_string, data = list(X=X, M=ncol(X), K=6, y = c(y,y2), N = nrow(X), iter=2000, warmup=1000,
                                                 alpha0=0.1), chains=3, cores=3, control=list(adapt_delta=0.9))
m2h
```

## 2 components
### medium overlap

generate new outcome

```{r}
y <- rnbinom(N1, size = 1.5, mu = exp(as.matrix(X) %*% c(1, -0.6, 0.4)))
y2 <- rnbinom(N2, size = 9, mu = exp(as.matrix(X) %*% c(1, 1.6, -0.2)))
```

plot again:

```{r}
df2 <- data.frame(val = c(y,y2), comp = c(rep(1, length(y)), rep(2, length(y2))))
ggplot(df2, aes(val, color=as.factor(comp))) + geom_density(adjust=1.8)
```

compute again. `stepFlexmix` is used when the model does not converge. Still the model for 4,5, and 6 components do not converge.

```{r comp2m, cache=T, warning=F}
comp2m1 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2), V1=X$V1, V2=X$V2), k = 1, model = FLXMRnegbin2())
comp2m2 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2), V1=X$V1, V2=X$V2), k = 2, model = FLXMRnegbin2())
comp2m3 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2), V1=X$V1, V2=X$V2), k = 3, model = FLXMRnegbin2())
comp2m4 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2), V1=X$V1, V2=X$V2), k = 4, model = FLXMRnegbin2())
comp2m5 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2), V1=X$V1, V2=X$V2), k = 5, model = FLXMRnegbin2())
comp2m6 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2), V1=X$V1, V2=X$V2), k = 6, model = FLXMRnegbin2())
```

```{r}
aics <- c(AIC(comp2m1), AIC(comp2m2), AIC(comp2m3))
bics <- c(BIC(comp2m1), BIC(comp2m2), BIC(comp2m3))
icls <- c(ICL(comp2m1), ICL(comp2m2), ICL(comp2m3))
```

```{r}
which.min(aics)
which.min(bics)
which.min(icls)
```

```{r stan2m, warning=F, cache=T}
m2m <- stan(model_code = model_string, data = list(X=X, M=ncol(X), K=6, y = c(y,y2), N = nrow(X), iter=2000, warmup=1000,
                                                 alpha0=0.1), chains=3, cores=3, control=list(adapt_delta=0.9))
m2m
```

## 2 components
## low overlap

generate new outcome

```{r}
y <- rnbinom(N1, size = 1.5, mu = exp(as.matrix(X) %*% c(1, -0.5, 0.3)))
y2 <- rnbinom(N2, size = 8, mu = exp(as.matrix(X) %*% c(1, 2.1, 0.2)))
```

plot the data

```{r }
df3 <- data.frame(val = c(y,y2), comp = c(rep(1, length(y)), rep(2, length(y2))))
ggplot(df3, aes(val, color=as.factor(comp))) + geom_density(adjust=1.8)
```

```{r , echo=F}
df2comp <- data.frame(val=c(df$val, df2$val, df3$val), comp = rep(df$comp, 3), 
                      overlap = c(rep("heavy", N1+N2),
                                  rep("medium", N1+N2),
                                  rep("low", N1+N2)))

write.csv(df2comp, "/home/ubu/ownCloud/docs/papers/paper1/data/2comp.csv", row.names = F)
```

compute models

```{r comp2l, cache=T, warning=F}
comp2l1 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2), V1=X$V1, V2=X$V2), k = 1, model = FLXMRnegbin2())
comp2l2 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2), V1=X$V1, V2=X$V2), k = 2, model = FLXMRnegbin2())
comp2l3 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2), V1=X$V1, V2=X$V2), k = 3, model = FLXMRnegbin2())
comp2l4 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2), V1=X$V1, V2=X$V2), k = 4, model = FLXMRnegbin2())
comp2l5 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2), V1=X$V1, V2=X$V2), k = 5, model = FLXMRnegbin2())
comp2l6 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2), V1=X$V1, V2=X$V2), k = 6, model = FLXMRnegbin2())
```

```{r }
aics <- c(AIC(comp2l1), AIC(comp2l2), AIC(comp2l3), AIC(comp2l6)) # 4, 5 does not converge
bics <- c(BIC(comp2l1), BIC(comp2l2), BIC(comp2l3), BIC(comp2l6)) # 4, 5 does not converge
icls <- c(ICL(comp2l1), ICL(comp2l2), ICL(comp2l3), ICL(comp2l6)) # 4, 5 does not converge
```

find lowest values

```{r }
which.min(aics)
which.min(bics)
which.min(icls)
```

```{r stan2l, warning=F, cache=T}
m2l <- stan(model_code = model_string, data = list(X=X, M=ncol(X), K=6, y = c(y,y2), N = nrow(X), iter=2000, warmup=1000,
                                                 alpha0=0.1), chains=3, cores=3, control=list(adapt_delta=0.9))
m2l
```

## 3 components
### heavy overlap

generate new data

```{r}
N1 <- 3000
N2 <- 5000
N3 <- 3000
X <- data.frame(V=rep(1,N1+N2+N3), V1=runif(N1+N2+N3, 0.8, 1), V2=runif(N1+N2+N3, 0.8, 1))
```

and outcome

```{r }
y <- rnbinom(N1, size = 1.2, mu = exp(as.matrix(X) %*% c(1, -0.5, 0.4)))
y2 <- rnbinom(N2, size = 10, mu = exp(as.matrix(X) %*% c(1, 1.0, -0.5)))
y3 <- rnbinom(N3, size = 14, mu = exp(as.matrix(X) %*% c(1, 1.2, -0.2)))
```

plot

```{r }
df <- data.frame(val = c(y,y2,y3), comp = c(rep(1, length(y)), rep(2, length(y2)), rep(3, length(y3))))
ggplot(df, aes(val, color=as.factor(comp))) + geom_density(adjust=3)
```

```{r comp3h, cache=T, warning=F}
comp3h1 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2,y3), V1=X$V1, V2=X$V2), k = 1, model = FLXMRnegbin2())
comp3h2 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2,y3), V1=X$V1, V2=X$V2), k = 2, model = FLXMRnegbin2())
comp3h3 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2,y3), V1=X$V1, V2=X$V2), k = 3, model = FLXMRnegbin2())
comp3h4 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2,y3), V1=X$V1, V2=X$V2), k = 4, model = FLXMRnegbin2())
comp3h5 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2,y3), V1=X$V1, V2=X$V2), k = 5, model = FLXMRnegbin2())
comp3h6 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2,y3), V1=X$V1, V2=X$V2), k = 6, model = FLXMRnegbin2())
```

```{r }
aics <- c(AIC(comp3h1), AIC(comp3h2), AIC(comp3h3), AIC(comp3h4), AIC(comp3h5), AIC(comp3h6))
bics <- c(BIC(comp3h1), BIC(comp3h2), BIC(comp3h3), BIC(comp3h4), BIC(comp3h5), BIC(comp3h6))
icls <- c(ICL(comp3h1), ICL(comp3h2), ICL(comp3h3), ICL(comp3h4), ICL(comp3h5), ICL(comp3h6))
```

```{r }
which.min(aics)
which.min(bics)
which.min(icls)
```

```{r stan3h, warning=F, cache=T}
m3h <- stan(model_code = model_string, data = list(X=X, M=ncol(X), K=6, y = c(y,y2,y3), N = nrow(X), iter=2000, warmup=1000,
                                                 alpha0=0.1), chains=3, cores=3, control=list(adapt_delta=0.9))
m3h
```

## 3 components
### medium overlap

generate outcome

```{r}
N1 <- 3000
N2 <- 5000
N3 <- 3000
y <- rnbinom(N1, size = 1.5, mu = exp(as.matrix(X) %*% c(1, -0.6, 0.4)))
y2 <- rnbinom(N2, size = 9, mu = exp(as.matrix(X) %*% c(1, 1.6, -0.1)))
y3 <- rnbinom(N3, size = 13, mu = exp(as.matrix(X) %*% c(1, 2.0, 0.2)))
```

plot

```{r }
df2 <- data.frame(val = c(y,y2,y3), comp = c(rep(1, length(y)), rep(2, length(y2)), rep(3, length(y3))))
ggplot(df2, aes(val, color=as.factor(comp))) + geom_density(adjust=3)
```

compute models

```{r comp3m, cache=T, warning=F}
comp3m1 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2,y3), V1=X$V1, V2=X$V2), k = 1, model = FLXMRnegbin2())
comp3m2 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2,y3), V1=X$V1, V2=X$V2), k = 2, model = FLXMRnegbin2())
comp3m3 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2,y3), V1=X$V1, V2=X$V2), k = 3, model = FLXMRnegbin2())
comp3m4 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2,y3), V1=X$V1, V2=X$V2), k = 4, model = FLXMRnegbin2())
comp3m5 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2,y3), V1=X$V1, V2=X$V2), k = 5, model = FLXMRnegbin2())
comp3m6 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2,y3), V1=X$V1, V2=X$V2), k = 6, model = FLXMRnegbin2())
```

```{r}
aics <- c(AIC(comp3m1), AIC(comp3m2), AIC(comp3m3), AIC(comp3m4), AIC(comp3m5), AIC(comp3m6))
bics <- c(BIC(comp3m1), BIC(comp3m2), BIC(comp3m3), BIC(comp3m4), BIC(comp3m5), BIC(comp3m6))
icls <- c(ICL(comp3m1), ICL(comp3m2), ICL(comp3m3), ICL(comp3m4), ICL(comp3m5), ICL(comp3m6))
```

```{r }
which.min(aics)
which.min(bics)
which.min(icls)
```

```{r stan3m, warning=F, cache=T}
m3m <- stan(model_code = model_string, data = list(X=X, M=ncol(X), K=6, y = c(y,y2,y3), N = nrow(X), iter=2000, warmup=1000,
                                                 alpha0=0.1), chains=3, cores=3, control=list(adapt_delta=0.9))
m3m
```

## 3 components
### low overlap

```{r }
N1 <- 3000
N2 <- 5000
N3 <- 3000
y <- rnbinom(N1, size = 1.5, mu = exp(as.matrix(X) %*% c(1, -0.5, 0.3)))
y2 <- rnbinom(N2, size = 8, mu = exp(as.matrix(X) %*% c(1, 2.1, 0.2)))
y3 <- rnbinom(N3, size = 12, mu = exp(as.matrix(X) %*% c(1, 3.1, 0.5)))
```

```{r}
df3 <- data.frame(val = c(y,y2,y3), comp = c(rep(1, length(y)), rep(2, length(y2)), rep(3, length(y3))))
ggplot(df3, aes(val, color=as.factor(comp))) + geom_density(adjust=2)
```

```{r }
df3comp <- data.frame(val=c(df$val, df2$val, df3$val), comp = rep(df$comp, 3), 
                      overlap = c(rep("heavy", N1+N2+N3),
                                  rep("medium", N1+N2+N3),
                                  rep("low", N1+N2+N3)))

write.csv(df3comp, "/home/ubu/ownCloud/docs/papers/paper1/data/3comp.csv", row.names = F)
```

```{r comp3l, cache=T, warning=F}
comp3l1 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2,y3), V1=X$V1, V2=X$V2), k = 1, model = FLXMRnegbin2())
comp3l2 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2,y3), V1=X$V1, V2=X$V2), k = 2, model = FLXMRnegbin2())
comp3l3 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2,y3), V1=X$V1, V2=X$V2), k = 3, model = FLXMRnegbin2())
comp3l4 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2,y3), V1=X$V1, V2=X$V2), k = 4, model = FLXMRnegbin2())
comp3l5 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2,y3), V1=X$V1, V2=X$V2), k = 5, model = FLXMRnegbin2())
comp3l6 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2,y3), V1=X$V1, V2=X$V2), k = 6, model = FLXMRnegbin2())
```

```{r}
aics <- c(AIC(comp3l1), AIC(comp3l2), AIC(comp3l3), AIC(comp3l4), AIC(comp3l5), AIC(comp3l6))
bics <- c(BIC(comp3l1), BIC(comp3l2), BIC(comp3l3), BIC(comp3l4), BIC(comp3l5), BIC(comp3l6))
icls <- c(ICL(comp3l1), ICL(comp3l2), ICL(comp3l3), ICL(comp3l4), ICL(comp3l5), ICL(comp3l6))
```

```{r }
which.min(aics)
which.min(bics)
which.min(icls)
```

```{r stan3l, warning=F, cache=T}
m3l <- stan(model_code = model_string, data = list(X=X, M=ncol(X), K=6, y = c(y,y2,y3), N = nrow(X), alpha0=0.1), 
            iter=2000, warmup=1000, chains=3, cores=3, control=list(adapt_delta=0.9))
m3l
```

## 4 components
### heavy overlap

```{r }
N1 <- 3000
N2 <- 5000
N3 <- 3000
N4 <- 4000
X <- data.frame(V=rep(1,N1+N2+N3+N4), V1=runif(N1+N2+N3+N4, 0.8, 1), V2=runif(N1+N2+N3+N4, 0.8, 1))
```

```{r}
y <- rnbinom(N1, size = 1.2, mu = exp(as.matrix(X) %*% c(1, -0.5, 0.4)))
y2 <- rnbinom(N2, size = 10, mu = exp(as.matrix(X) %*% c(1, 1.0, -0.5)))
y3 <- rnbinom(N3, size = 14, mu = exp(as.matrix(X) %*% c(1, 1.2, -0.2)))
y4 <- rnbinom(N4, size = 17, mu = exp(as.matrix(X) %*% c(1, 1.6, 0.1)))
```

```{r }
df <- data.frame(val = c(y,y2,y3,y4), comp = c(rep(1, length(y)), rep(2, length(y2)), rep(3, length(y3)), rep(4, length(y4))))
ggplot(df, aes(val, color=as.factor(comp))) + geom_density(adjust=4)
```

```{r comp4h, cache=T, warning=F }
comp4h1 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2,y3,y4), V1=X$V1, V2=X$V2), k = 1, model = FLXMRnegbin2())
comp4h2 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2,y3,y4), V1=X$V1, V2=X$V2), k = 2, model = FLXMRnegbin2())
comp4h3 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2,y3,y4), V1=X$V1, V2=X$V2), k = 3, model = FLXMRnegbin2())
comp4h4 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2,y3,y4), V1=X$V1, V2=X$V2), k = 4, model = FLXMRnegbin2())
comp4h5 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2,y3,y4), V1=X$V1, V2=X$V2), k = 5, model = FLXMRnegbin2())
comp4h6 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2,y3,y4), V1=X$V1, V2=X$V2), k = 6, model = FLXMRnegbin2())
```

```{r}
aics <- c(AIC(comp4h1), AIC(comp4h2), AIC(comp4h3), AIC(comp4h4), AIC(comp4h5), AIC(comp4h6))
bics <- c(BIC(comp4h1), BIC(comp4h2), BIC(comp4h3), BIC(comp4h4), BIC(comp4h5), BIC(comp4h6))
icls <- c(ICL(comp4h1), ICL(comp4h2), ICL(comp4h3), ICL(comp4h4), ICL(comp4h5), ICL(comp4h6))
```

```{r }
which.min(aics)
which.min(bics)
which.min(icls)
```

```{r stan4h, warning=F, cache=T}
m4h <- stan(model_code = model_string, data = list(X=X, M=ncol(X), K=6, y = c(y,y2,y3,y4), N = nrow(X), alpha0=0.1), 
            iter=2000, warmup=1000, chains=3, cores=3, control=list(adapt_delta=0.9))
m4h
```

## 4 components
## medium overlap

```{r }
N1 <- 3000
N2 <- 5000
N3 <- 3000
N4 <- 4000
y <- rnbinom(N1, size = 1.5, mu = exp(as.matrix(X) %*% c(1, -0.6, 0.4)))
y2 <- rnbinom(N2, size = 10, mu = exp(as.matrix(X) %*% c(1, 1.6, -0.2)))
y3 <- rnbinom(N3, size = 13, mu = exp(as.matrix(X) %*% c(1, 2.0, 0.2)))
y4 <- rnbinom(N4, size = 15, mu = exp(as.matrix(X) %*% c(1, 2.5, 0.3)))
```

```{r}
df2 <- data.frame(val = c(y,y2,y3,y4), comp = c(rep(1, length(y)), rep(2, length(y2)), rep(3, length(y3)), rep(4, length(y4))))
ggplot(df2, aes(val, color=as.factor(comp))) + geom_density(adjust=3)
```

```{r comp4m, cache=T, warning=F}
comp4m1 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2,y3,y4), V1=X$V1, V2=X$V2), k = 1, model = FLXMRnegbin2())
comp4m2 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2,y3,y4), V1=X$V1, V2=X$V2), k = 2, model = FLXMRnegbin2())
comp4m3 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2,y3,y4), V1=X$V1, V2=X$V2), k = 3, model = FLXMRnegbin2())
comp4m4 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2,y3,y4), V1=X$V1, V2=X$V2), k = 4, model = FLXMRnegbin2())
comp4m5 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2,y3,y4), V1=X$V1, V2=X$V2), k = 5, model = FLXMRnegbin2())
comp4m6 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2,y3,y4), V1=X$V1, V2=X$V2), k = 6, model = FLXMRnegbin2())

```

```{r }
aics <- c(AIC(comp4m1), AIC(comp4m2), AIC(comp4m3), AIC(comp4m4), AIC(comp4m5), AIC(comp4m6))
bics <- c(BIC(comp4m1), BIC(comp4m2), BIC(comp4m3), BIC(comp4m4), BIC(comp4m5), BIC(comp4m6))
icls <- c(ICL(comp4m1), ICL(comp4m2), ICL(comp4m3), ICL(comp4m4), ICL(comp4m5), ICL(comp4m6))
```

```{r}
which.min(aics)
which.min(bics)
which.min(icls)
```

```{r stan4m, warning=F, cache=T}
m4m <- stan(model_code = model_string, data = list(X=X, M=ncol(X), K=6, y = c(y,y2,y3,y4), N = nrow(X), alpha0=0.1), 
            iter=2000, warmup=1000, chains=3, cores=3, control=list(adapt_delta=0.9))
m4m
```


## 4 components
## low overlap

```{r }
N1 <- 3000
N2 <- 5000
N3 <- 3000
N4 <- 4000
y <- rnbinom(N1, size = 1.5, mu = exp(as.matrix(X) %*% c(1, -0.5, 0.3)))
y2 <- rnbinom(N2, size = 8, mu = exp(as.matrix(X) %*% c(1, 2.1, 0.2)))
y3 <- rnbinom(N3, size = 12, mu = exp(as.matrix(X) %*% c(1, 3.1, 0.5)))
y4 <- rnbinom(N4, size = 18, mu = exp(as.matrix(X) %*% c(1, 4.0, 0.2)))
```

```{r }
df3 <- data.frame(val = c(y,y2,y3,y4), comp = c(rep(1, length(y)), rep(2, length(y2)), rep(3, length(y3)), rep(4, length(y4))))
ggplot(df3, aes(val, color=as.factor(comp))) + geom_density(adjust=3)
```

```{r, echo=F}
df4comp <- data.frame(val=c(df$val, df2$val, df3$val), comp = rep(df$comp, 3), 
                      overlap = c(rep("heavy", N1+N2+N3+N4),
                                  rep("medium", N1+N2+N3+N4),
                                  rep("low", N1+N2+N3+N4)))

write.csv(df4comp, "/home/ubu/ownCloud/docs/papers/paper1/data/4comp.csv", row.names = F)
```

```{r comp4l, cache=T, warning=F}
comp4l1 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2,y3,y4), V1=X$V1, V2=X$V2), k = 1, model = FLXMRnegbin2())
comp4l2 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2,y3,y4), V1=X$V1, V2=X$V2), k = 2, model = FLXMRnegbin2())
comp4l3 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2,y3,y4), V1=X$V1, V2=X$V2), k = 3, model = FLXMRnegbin2())
comp4l4 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2,y3,y4), V1=X$V1, V2=X$V2), k = 4, model = FLXMRnegbin2())
comp4l5 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2,y3,y4), V1=X$V1, V2=X$V2), k = 5, model = FLXMRnegbin2())
comp4l6 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2,y3,y4), V1=X$V1, V2=X$V2), k = 6, model = FLXMRnegbin2())
```

```{r }
aics <- c(AIC(comp4l1), AIC(comp4l2), AIC(comp4l3), AIC(comp4l4), AIC(comp4l5), AIC(comp4l6))
bics <- c(BIC(comp4l1), BIC(comp4l2), BIC(comp4l3), BIC(comp4l4), BIC(comp4l5), BIC(comp4l6))
icls <- c(ICL(comp4l1), ICL(comp4l2), ICL(comp4l3), ICL(comp4l4), ICL(comp4l5), ICL(comp4l6))
```

```{r}
which.min(aics)
which.min(bics)
which.min(icls)
```

```{r stan4l, warning=F, cache=T}
m4l <- stan(model_code = model_string, data = list(X=X, M=ncol(X), K=6, y = c(y,y2,y3,y4), N = nrow(X), alpha0=0.1), 
            iter=2000, warmup=1000, chains=3, cores=3, control=list(adapt_delta=0.9))
m4l
```


## 5 components
### heavy overlap

```{r }
N1 <- 3000
N2 <- 5000
N3 <- 3000
N4 <- 4000
N5 <- 2000
X <- data.frame(V=rep(1,N1+N2+N3+N4+N5), V1=runif(N1+N2+N3+N4+N5, 0.8, 1), V2=runif(N1+N2+N3+N4+N5, 0.8, 1))
```

```{r }
y <- rnbinom(N1, size = 1.2, mu = exp(as.matrix(X) %*% c(1, -0.5, 0.4)))
y2 <- rnbinom(N2, size = 10, mu = exp(as.matrix(X) %*% c(1, 1.0, -0.5)))
y3 <- rnbinom(N3, size = 14, mu = exp(as.matrix(X) %*% c(1, 1.2, -0.2)))
y4 <- rnbinom(N4, size = 17, mu = exp(as.matrix(X) %*% c(1, 1.6, 0.1)))
y5 <- rnbinom(N5, size = 19, mu = exp(as.matrix(X) %*% c(1, 2.1, 0.2)))
```

```{r}
df <- data.frame(val = c(y,y2,y3,y4,y5), comp = c(rep(1, length(y)), rep(2, length(y2)), rep(3, length(y3)), rep(4, length(y4)), rep(5, length(y5)) ))
ggplot(df, aes(val, color=as.factor(comp))) + geom_density(adjust=3)
```

```{r comp5h, cache=T, warning=F}
comp5h1 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2,y3,y4,y5), V1=X$V1, V2=X$V2), k = 1, model = FLXMRnegbin2())
comp5h2 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2,y3,y4,y5), V1=X$V1, V2=X$V2), k = 2, model = FLXMRnegbin2())
comp5h3 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2,y3,y4,y5), V1=X$V1, V2=X$V2), k = 3, model = FLXMRnegbin2())
comp5h4 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2,y3,y4,y5), V1=X$V1, V2=X$V2), k = 4, model = FLXMRnegbin2())
comp5h5 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2,y3,y4,y5), V1=X$V1, V2=X$V2), k = 5, model = FLXMRnegbin2())
comp5h6 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2,y3,y4,y5), V1=X$V1, V2=X$V2), k = 6, model = FLXMRnegbin2())
```

```{r }
aics <- c(AIC(comp5h1), AIC(comp5h2), AIC(comp5h3), AIC(comp5h4), AIC(comp5h5), AIC(comp5h6))
bics <- c(BIC(comp5h1), BIC(comp5h2), BIC(comp5h3), BIC(comp5h4), BIC(comp5h5), BIC(comp5h6))
icls <- c(ICL(comp5h1), ICL(comp5h2), ICL(comp5h3), ICL(comp5h4), ICL(comp5h5), ICL(comp5h6))
```

```{r}
which.min(aics)
which.min(bics)
which.min(icls)
```

```{r stan5h, warning=F, cache=T}
m5h <- stan(model_code = model_string, data = list(X=X, M=ncol(X), K=7, y = c(y,y2,y3,y4,y5), N = nrow(X), alpha0=0.1), 
            iter=2000, warmup=1000, chains=3, cores=3, control=list(adapt_delta=0.9))
m5h
```


## 5 components
## medium overlap

```{r }
N1 <- 3000
N2 <- 5000
N3 <- 3000
N4 <- 4000
N5 <- 2000
y <- rnbinom(N1, size = 1.5, mu = exp(as.matrix(X) %*% c(1, -0.6, 0.4)))
y2 <- rnbinom(N2, size = 10, mu = exp(as.matrix(X) %*% c(1, 1.6, -0.2)))
y3 <- rnbinom(N3, size = 13, mu = exp(as.matrix(X) %*% c(1, 2.0, 0.2)))
y4 <- rnbinom(N4, size = 15, mu = exp(as.matrix(X) %*% c(1, 2.5, 0.3)))
y5 <- rnbinom(N5, size = 22, mu = exp(as.matrix(X) %*% c(1, 3.4, 0.4)))
```

```{r }
df2 <- data.frame(val = c(y,y2,y3,y4,y5), comp = c(rep(1, length(y)), rep(2, length(y2)), rep(3, length(y3)), rep(4, length(y4)), rep(5, length(y5)) ))
ggplot(df2, aes(val, color=as.factor(comp))) + geom_density(adjust=3)
```

```{r comp5m, cache=T, warning=F}
comp5m1 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2,y3,y4,y5), V1=X$V1, V2=X$V2), k = 1, model = FLXMRnegbin2())
comp5m2 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2,y3,y4,y5), V1=X$V1, V2=X$V2), k = 2, model = FLXMRnegbin2())
comp5m3 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2,y3,y4,y5), V1=X$V1, V2=X$V2), k = 3, model = FLXMRnegbin2())
comp5m4 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2,y3,y4,y5), V1=X$V1, V2=X$V2), k = 4, model = FLXMRnegbin2())
comp5m5 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2,y3,y4,y5), V1=X$V1, V2=X$V2), k = 5, model = FLXMRnegbin2())
comp5m6 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2,y3,y4,y5), V1=X$V1, V2=X$V2), k = 6, model = FLXMRnegbin2())
```

```{r }
aics <- c(AIC(comp5m1), AIC(comp5m2), AIC(comp5m3), AIC(comp5m4), AIC(comp5m5), AIC(comp5m6))
bics <- c(BIC(comp5m1), BIC(comp5m2), BIC(comp5m3), BIC(comp5m4), BIC(comp5m5), BIC(comp5m6))
icls <- c(ICL(comp5m1), ICL(comp5m2), ICL(comp5m3), ICL(comp5m4), ICL(comp5m5), ICL(comp5m6))
```

```{r }
which.min(aics)
which.min(bics)
which.min(icls)
```

```{r stan5m, warning=F, cache=T}
m5m <- stan(model_code = model_string, data = list(X=X, M=ncol(X), K=7, y = c(y,y2,y3,y4,y5), N = nrow(X), alpha0=0.1), 
            iter=2000, warmup=1000, chains=3, cores=3, control=list(adapt_delta=0.9))
m5m
```

## 5 components
## low overlap

```{r}
N1 <- 3000
N2 <- 5000
N3 <- 3000
N4 <- 4000
N5 <- 2000
y <- rnbinom(N1, size = 1.5, mu = exp(as.matrix(X) %*% c(1, -0.6, 0.4)))
y2 <- rnbinom(N2, size = 7, mu = exp(as.matrix(X) %*% c(1, 2.2, 0.5)))
y3 <- rnbinom(N3, size = 14, mu = exp(as.matrix(X) %*% c(1, 2.8, 0.8)))
y4 <- rnbinom(N4, size = 19, mu = exp(as.matrix(X) %*% c(1, 4.0, 0.4)))
y5 <- rnbinom(N5, size = 23, mu = exp(as.matrix(X) %*% c(1, 4.7, 0.3)))
```

```{r }
df3 <- data.frame(val = c(y,y2,y3,y4,y5), comp = c(rep(1, length(y)), rep(2, length(y2)), rep(3, length(y3)), rep(4, length(y4)), rep(5, length(y5)) ))
ggplot(df3, aes(val, color=as.factor(comp))) + geom_density(adjust=3)
```

```{r ,echo=F}
df5comp <- data.frame(val=c(df$val, df2$val, df3$val), comp = rep(df$comp, 3), 
                      overlap = c(rep("heavy", N1+N2+N3+N4+N5),
                                  rep("medium", N1+N2+N3+N4+N5),
                                  rep("low", N1+N2+N3+N4+N5)))

write.csv(df5comp, "/home/ubu/ownCloud/docs/papers/paper1/data/5comp.csv", row.names = F)
```

```{r comp5l, cache=T, warning=F}
comp5l1 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2,y3,y4,y5), V1=X$V1, V2=X$V2), k = 1, model = FLXMRnegbin2())
comp5l2 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2,y3,y4,y5), V1=X$V1, V2=X$V2), k = 2, model = FLXMRnegbin2())
comp5l3 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2,y3,y4,y5), V1=X$V1, V2=X$V2), k = 3, model = FLXMRnegbin2())
comp5l4 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2,y3,y4,y5), V1=X$V1, V2=X$V2), k = 4, model = FLXMRnegbin2())
comp5l5 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2,y3,y4,y5), V1=X$V1, V2=X$V2), k = 5, model = FLXMRnegbin2())
comp5l6 <- flexmix(y ~ V1+V2, data = data.frame(y=c(y,y2,y3,y4,y5), V1=X$V1, V2=X$V2), k = 6, model = FLXMRnegbin2())
```

```{r }
aics <- c(AIC(comp5l1), AIC(comp5l2), AIC(comp5l3), AIC(comp5l4), AIC(comp5l5), AIC(comp5l6))
bics <- c(BIC(comp5l1), BIC(comp5l2), BIC(comp5l3), BIC(comp5l4), BIC(comp5l5), BIC(comp5l6))
icls <- c(ICL(comp5l1), ICL(comp5l2), ICL(comp5l3), ICL(comp5l4), ICL(comp5l5), ICL(comp5l6))
```

```{r }
which.min(aics)
which.min(bics)
which.min(icls)
```

```{r stan5l, warning=F, cache=T}
m5l <- stan(model_code = model_string, data = list(X=X, M=ncol(X), K=7, y = c(y,y2,y3,y4,y5), N = nrow(X), alpha0=0.1), 
            iter=2000, warmup=1000, chains=3, cores=3, control=list(adapt_delta=0.9))
m5l
```
